\begin{acknowledgements}
I'd like to express my deep gratitude to my professors, friends and family for putting up with me during the last few months of completing this work.
This endeavor has been a great exercise in reconciling my idealistic vision of perfection with the constraints imposed by finite time and energy.
I'm grateful to my mentors for having allowed me enough creative freedom in this process to explore the concepts and techniques that most interested me at the time.
I've learned so much both about myself and about the process of working effectively as an applied mathematician over the course of the process of writing this thesis.
To name some topics that come to my mind: mathematical modeling, numerical methods, Fortran, debugging, profiling, parallel computing, HPC resource orchestration, version control, build automation, verification of codes and calculations, routine, sleep, diet, time management, interpersonal communication, the importance of social interaction, and deciding when to stop working.
Without the particular challenges and opportunities afforded me during this project, I may have avoided many uncomfortable and important lessons I've learned lately.

In particular, in addition to my advisors, Dr. Kevin Kreider and Dr. Curtis Clemons, I'd like to thank Dr. Shane Rogers at Clarkson University in Potsdam, New York and Dr. Ole Jacob Broch at SINTEF Ocean in Trondheim, Norway, with whom I began this project in Summer 2016.
Also, thanks to Dr. Jutta Luettmer--Strathmann for staying up late with me one evening in the physics department helping me to prepare for my defense.
The ongoing support of these people and many others has been invaluable, in practical advice and especially in emotional encouragement.

I would also like to express my gratitude to the kind folks at the NSF Pacific Research Platform including my uncle, John Graham and his colleague Dima Mishin for providing me access to and user support for Nautilus, a distributed, heterogeneity's HPC cluster which I used for many thousands of CPU hours to run simulations for this thesis.

And finally, thanks to my parents for feeding and housing me at the beginning and end of my studies.
I feel extraordinarily blessed to have been born into a stable, loving family that has never stopped urging me forward.

\vfill

This project was supported in part by the US National Science
Foundation under Grant No. EEC--1359256, and by the Norwegian National Research
Council, Project number 254883/E40.
The Nautilus computational cluster on which the simulations in this thesis were run is supported by NSF awards 
CNS 0821155, CNS-1338192, CNS-1456638, CNS-1730158, 
ACI-1540112, and ACI-1541349

\end{acknowledgements}






